\Huge Don't interpolate your data!

### 2022-11-03 paper/main.tex
merge changes with hogg main

Is it astronomical blasphemy to say that this practice is \emph{wrong?}
It is wrong because minimum-variance unbiased estimators are (often) maximum-likelihood estimators (\citealt{something}).
The deep, high signal-to-noise image or spectrum we seek is a statistical estimate of the image or spectrum, given noisy data.
Noisy data are combined to make a measurement by building and optimizing a likelihood, not by shifting and stacking!
Of course we don't always think of an image or a spectrum of the sky as the result of applying an \emph{estimator}.
But it is.
It is just a simultaneous estimate of a lot of different components (a lot of different pixels; all the pixels in the combined image or spectrum).

Our view is that image combination should be seen as a problem in statistical estimation.
Let's find the best estimator for the combined image we can find, and let's deliver that (along with an uncertainty estimate).
The data are never manipulated when using an estimator, but the relationship between the per epoch observations (inputs) and `combined image' or `combined spectrum' (outputs) is identical to the input--output relationship of a shift-and-coadd approach.
Unlike the shift-and-coadd approach, an estimator is insensitive to under- or poorly-sampled data (e.g., \textsl{APOGEE} raw data).
As we will show, an estimator also yields a more faithful representation\footnote{
    An estimator needs a \emph{model} to produce a combined spectrum, but we use the word \emph{representation} to avoid overloading the word `model', and to indicate that we will not construct a physical model of a spectrum from atomic physics.
    We only need a flexible representation!
}
of the underlying \emph{true} spectrum, critically important for any subsequent precision measurements (e.g., radial velocities, chemical abundances).

Things to emphasize here and below:
\begin{itemize}
\item The input--output relationship of this modeling approach is identical to the input--output relationship of the shift-and-coadd approach. It is just better.
\item People have a lot of hangups about the word ``model''. We have to use another word, or be very very careful how we use this word. For instance, some people I talked to about this thought I meant a physical model of the spectrum starting with atomic physics!
\item We have to explain why and how we are better than other approaches in terms that people care about (like radial-velocity precision and abundances etc).
\item All this is even more important in ill-sampled contexts than well-sampled contexts. See, for example, \textsl{APOGEE} raw data!
\item We should choose a rigid terminology and stick to it here. HOGG is currently liking ``combined spectrum'' for the model expectation.
\item This model is a sub-model of \textsl{wobble} \cite{wobble}.
\end{itemize}

>>>>

Is it astronomical blasphemy to say that this practice is \emph{wrong?}
It is wrong because minimum-variance unbiased estimators require care in their construction; they are (often) maximum-likelihood estimators (\citealt{something}).
The deep, high signal-to-noise image or spectrum we seek is a statistical estimate of the image or spectrum, given the noisy data of the individual exposures.
Noisy data are combined to make a measurement by building and optimizing a likelihood, not by shifting and stacking the input data.
Of course we don't always think of an image or a spectrum of the sky as the result of applying an \emph{estimator},
but it is.
It is just a simultaneous estimate of a lot of different components (a lot of different pixels; all the pixels in the combined image or spectrum).

We are in a period in astrophysics in which exceedingly precise measurements are expected, in both imaging and spectroscopy projects.
In spectroscopy, in particular, high signal-to-noise spectral measurements are expected to obtain surface-abundance measurements with precisions a few percent, and radial-velocity measurements with precisions of up to one ten-thousandth of a \emph{pixel} (in terms of the Doppler shift).
Spectral combination methods that distort absorption (or emission) lines will translate into biases, errors, or increased variance in these measurements.
Spectral combination methods that correlate the noise in the outputs will translate into increased variance, and necessitate more sophisticated downstream measurement techniques (which don't really exist at present).

HOGG: Of course if your GOAL IS RV you should DO SOMETHING DIFFERENT.

One of the great things about the standard practice---the simple process of shifting and stacking the data---is that this procedure makes almost no assumptions about what the output, combined image or spectrum will look like.
It isn't really a ``model'' for the data; it is more like a ``summary'' of the data.
When we switch to a statistical-inference framework, we have to choose some form or basis for expressing the combined image.
In what follows we use a maximally parameterized model, in the sense that we choose an expression for the functions or function space that has as many parameters as the output image has pixels.
That means that the approach we take does not get its power from the choice of parameterized functions for the output:
Indeed it has just as much freedom as the shifting-and-stacking standard practice.
The forward-modeling we recommend gets its power from the fact that it generates the data directly and without manipulation.

The data are never manipulated when using an estimator, but the relationship between the per epoch observations (inputs) and `combined image' or `combined spectrum' (outputs) is identical to the input--output relationship of a shift-and-coadd approach.
Unlike the shift-and-coadd approach, an estimator is insensitive to under- or poorly-sampled data (e.g., \textsl{APOGEE} raw data).
As we will show, an estimator also yields a more faithful representation\footnote{
    An estimator needs a \emph{model} to produce a combined spectrum, but we use the word \emph{representation} to avoid overloading the word `model', and to indicate that we will not construct a physical model of a spectrum from atomic physics.
    We only need a flexible representation!
}
of the underlying \emph{true} spectrum, critically important for any subsequent precision measurements (e.g., radial velocities, chemical abundances).

Things to emphasize here and below:
\begin{itemize}
    \item Maybe move this to the method section? People have a lot of hangups about the word ``model''. We have to use another word, or be very very careful how we use this word. For instance, some people I talked to about this thought I meant a physical model of the spectrum starting with atomic physics!
    \item All this is even more important in ill-sampled contexts than well-sampled contexts. See, for example, \textsl{APOGEE} raw data!
\item We should choose a rigid terminology and stick to it here. HOGG is currently liking ``combined spectrum'' for the model expectation.
\item This problem is a sub-problem of \textsl{wobble} \cite{wobble}.
\item The problem we solve here is the easy case. In the hard case, there is a slit, and the object is extended, and different observations are at different slit angles, different slit positions, and with different atmospheric seeing. That problem is a hard problem! But it probably solves the same way. Say a bit here and more in the discussion.
\end{itemize}
